MSCA 31006: Final Project - AQI Forecasting

### 2023-08-16 model/SARIMA.Rmd
Rename MSCA_31006_Time_Series_Final_Project - ARIMA.Rmd to SARIMA.Rmd

# Fit ARIMA model using auto.arima()
# Find appropriate values for d and D using ndiffs and nsdiffs
# d <- ndiffs(train_weekly_diff)
# D <- nsdiffs(train_weekly_diff)

# Fit ARIMA model using auto.arima() with seasonal components and previously found d and D
arima_model_weekly <- auto.arima(train_weekly, seasonal = TRUE, stepwise = FALSE,
                                 approximation = FALSE, trace = FALSE, lambda = "auto",
                                 start.p = 1, max.p = 4,
                                 start.q = 2, max.q = 5, 
                                 d=0)

@@

```
```{r}
forecast_arima_weekly
```

@@

#### ARIMA model using auto.arima() - Monthly

```{r}
# Convert monthly data to a time series object
# pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999, 1), frequency = 12)

y_values_monthly <- train_monthly
optimal_lambda <- BoxCox.lambda(y_values_monthly)

# Apply the Box-Cox transformation
transformed_data_monthly <- if (optimal_lambda != 0) {
  (y_values_monthly^optimal_lambda - 1) / optimal_lambda
} else {
  log(y_values_monthly)
}

# Check stationarity, perform differencing if needed
kpss_test_result <- kpss.test(pm2_5_monthly_ts)
adf_test_result <- adf.test(pm2_5_monthly_ts)

if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
  pm2_5_monthly_diff <- diff(pm2_5_monthly_ts)
} else {
  pm2_5_monthly_diff <- pm2_5_monthly_ts
}

# Check stationarity, perform differencing if needed
kpss_test_result <- kpss.test(transformed_data_monthly)
adf_test_result <- adf.test(transformed_data_monthly)

if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
  pm2_5_monthly_diff <- diff(transformed_data_monthly)

  # Check stationarity after first differencing
  kpss_test_result_diff <- kpss.test(pm2_5_monthly_diff)
  adf_test_result_diff <- adf.test(pm2_5_monthly_diff)

  if (kpss_test_result_diff$p.value < 0.05 || adf_test_result_diff$p.value < 0.05) {
    # If still non-stationary, apply second-order differencing
    pm2_5_monthly_diff <- diff(pm2_5_monthly_diff)

    print("First and Second Order Differencing was performed")
  }
} else {
  pm2_5_monthly_diff <- transformed_data_monthly
  print("First Order Differencing was performed")
}

# Fit ARIMA model using auto.arima()
arima_model_monthly <- auto.arima(pm2_5_monthly_diff, seasonal = TRUE)

# Forecast using ARIMA model
h <- length(test_monthly)
forecast_arima_monthly <- forecast(arima_model_monthly, h = h)

original_scale_forecasts_monthly <- InvBoxCox(forecast_arima_monthly$mean, lambda = optimal_lambda)

print(cat("Optimal Lambda Value:", optimal_lambda, "\n"))

# Plot the original data
original_scale_forecasts_monthly <- original_scale_forecasts_monthly * 100
plot(train_monthly, main = "Original Data and Forecasts")
lines(original_scale_forecasts_monthly, col = "blue")  # Add forecast line in blue

# Add legend
legend("topleft", legend = c("Original Data", "Forecasts"), col = c("black", "blue"), lwd = 1)

```
```{r}

checkresiduals(arima_model_monthly)
#Ljung-Box test .05179

plot(pm2_5_monthly_diff, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_diff), max(pm2_5_monthly_diff)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
lines(test_monthly, col="red") # Test data in red
lines(original_scale_forecasts_monthly, col="blue", type="o") # Forecasted data in blue
legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))

mae <- mean(abs(test_monthly - original_scale_forecasts_monthly))
print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
#Mean Absolute Error (MAE): 9.31

comparison_df <- data.frame(
  Date = time(test_monthly),
  Actual = as.vector(test_monthly),
  Forecasted = as.vector(original_scale_forecasts_monthly)
)
print(comparison_df)

>>>>

# Fit ARIMA model using auto.arima()
arima_model_weekly <- auto.arima(train_weekly, seasonal = TRUE, stepwise = FALSE,
                                 trace = FALSE,
                                 start.p = 1, max.p = 4,
                                 start.q = 2, max.q = 5)

@@

@@
